---
title: "Manuscript statistics — full output"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
classoption: landscape
geometry: margin=0.5in
header-includes:
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,breakanywhere,fontsize=\scriptsize,commandchars=\\\{\}}
  - \makeatletter\def\verbatim@font{\scriptsize\ttfamily}\makeatother
---

<!--
Analysis code below is copied verbatim from "2026_10_06 manuscript_statistics.R".
The only additions are (1) the "knitting setup" chunk, which changes how output is
displayed so nothing is truncated in the PDF, and (2) an optional export chunk at the end.

Save this .Rmd in the same folder you normally run the .R script from (the one that
contains "Measures/", "DEG Outputs/", "Function Scripts/", etc.), or set root.dir below.
-->

```{r knitting-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE, error = TRUE,
                      comment = "", dev = "png", dpi = 150,
                      fig.width = 7, fig.height = 4.5)
# REQUIRED unless this .Rmd sits in your project folder: the folder that contains "Measures/", "DEG Outputs/", etc.
knitr::opts_knit$set(root.dir = "/Users/ggraham/Desktop/multiome_poa")

# Display only: print every row and every column of every table.
# Tibbles are printed as data.frames so wide tables wrap into column blocks
# instead of running off the PDF page. Values are not changed.
options(width = 150, digits = 7, max.print = 1e6,
        tibble.print_max = Inf, tibble.print_min = Inf, tibble.width = Inf)
registerS3method("print", "tbl_df", function(x, ...) {
  df <- as.data.frame(x)
  if (nrow(df) == 0) cat("<0 rows>\n") else print.data.frame(df, row.names = FALSE)
  invisible(x)
})
```

```{r input-check, error=FALSE}
# Stops the knit right away if the working directory is wrong, instead of producing a partial PDF.
cat("Working directory:", getwd(), "\n")
inputs <- c("Function Scripts/Dependencies/Term2gene_clown_go2.rds",
            "Function Scripts/Dependencies/Term2name.rds",
            "Measures/2025_12_26 all_data.csv",
            "DEG Outputs/FINAL degs classified w singular.csv",
            "Collaboration/all_clusters_DARs_peak_level_classified_with_support.csv",
            "Manuscript/Supplementary Tables/classifier_scores.csv",
            "Manuscript/updatedcluster_6_steroid_receptor_SPECIFIC_PROMOTER_ZSCORES.csv",
            paste0("DEG Outputs/05_11_2025 Neg Bin w Doms New_clustering/cluster_", c(0, 1, 6), ".csv"),
            "~/Desktop/optimal_clustering_rna_only.rds")
print(data.frame(file = inputs, found = file.exists(inputs)), row.names = FALSE)
if (!all(file.exists(inputs))) stop("Missing input files: set root.dir in the knitting-setup chunk to your project folder.")
```

# Setup

```{r setup-analysis}
#### Setup ####
library(Seurat)
library(tidyverse)
library(emmeans)
library(car)
library(lme4)
library(CytoTRACE)
library(AUCell)
set.seed(0)
options(pillar.sigfig = 5)

status_to_phase <- c(M = "M", D = "I", S = "IP", E = "LI", EP = "LIP",
                     F = "F", NF = "NF", NM = "NM", NRM = "NRM")
mif <- c("M", "I", "F")
all_phases <- c("M", "I", "IP", "LI", "LIP", "F", "NF", "NM")

add_phase <- function(d, keep = mif) {
  d %>%
    mutate(Phase = unname(status_to_phase[as.character(Status)])) %>%
    filter(Phase %in% keep) %>%
    mutate(Phase = factor(Phase, levels = keep))
}

model_label <- function(fit) {
  fn <- if (inherits(fit, "merMod")) "lmer" else if (inherits(fit, "glm")) "glm" else "lm"
  paste0(fn, "(", deparse1(formula(fit)), if (fn == "glm") ", family = binomial", ")")
}

# Table 1 format: Type-III test per term, uncorrected emmeans pairwise p-values on the grouping term
# contrasts = NULL returns every pairwise contrast
stat_table <- function(models, figure, group = "Phase", contrasts = c("M-F", "M-I", "I-F")) {
  imap_dfr(models, function(fit, trait) {
    a <- as.data.frame(Anova(fit, type = 3))
    a <- a[!rownames(a) %in% c("(Intercept)", "Residuals"), , drop = FALSE]
    stat <- intersect(c("F value", "LR Chisq", "Chisq"), names(a))[1]
    pw <- as.data.frame(pairs(emmeans(fit, group), adjust = "none"))
    lab <- gsub(" ", "", as.character(pw$contrast))
    want <- if (is.null(contrasts)) lab else contrasts
    pw_p <- map_dbl(want, function(x) {
      i <- which(lab %in% c(x, paste(rev(strsplit(x, "-")[[1]]), collapse = "-")))
      if (length(i)) pw$p.value[i[1]] else NA_real_
    })
    out <- tibble(Trait = trait, Model = model_label(fit), Figure = figure, Term = rownames(a),
                  !!stat := a[[stat]], df = a$Df, `Type-III p-value` = a[[ncol(a)]])
    if (!inherits(fit, "glm") && inherits(fit, "lm")) out$df_resid <- df.residual(fit)
    for (k in seq_along(want)) out[[want[k]]] <- ifelse(out$Term == group, pw_p[k], NA_real_)
    out
  })
}

# Raw group means; fold = mean2 / mean1, pct_diff = % change from group1 to group2
# NA when the trait is in no_ratio or either mean is <= 0
fold_table <- function(data, traits, group = "Phase", no_ratio = character()) {
  map_dfr(traits, function(v) {
    m <- data %>%
      filter(!is.na(.data[[v]]), !is.na(.data[[group]])) %>%
      group_by(level = .data[[group]]) %>%
      summarize(mean = mean(.data[[v]]), .groups = "drop")
    if (nrow(m) < 2) return(NULL)
    ij <- combn(nrow(m), 2)
    tibble(trait = v,
           group1 = as.character(m$level[ij[1, ]]), group2 = as.character(m$level[ij[2, ]]),
           mean1 = m$mean[ij[1, ]], mean2 = m$mean[ij[2, ]]) %>%
      mutate(difference = mean2 - mean1,
             ok = !(v %in% no_ratio) & mean1 > 0 & mean2 > 0,
             fold = ifelse(ok, mean2 / mean1, NA_real_),
             pct_diff = ifelse(ok, 100 * (mean2 - mean1) / mean1, NA_real_)) %>%
      dplyr::select(-ok)
  })
}

mean_se <- function(data, traits, group = "Phase") {
  data %>%
    dplyr::select(all_of(c(group, traits))) %>%
    pivot_longer(-all_of(group), names_to = "trait") %>%
    filter(!is.na(value)) %>%
    group_by(trait, .data[[group]]) %>%
    summarize(mean = mean(value), se = sd(value) / sqrt(n()), n = n(), .groups = "drop")
}

# Per-individual cell counts in each cluster
cluster_counts <- function(o, cluster_col) {
  o@meta.data %>%
    count(individual, Status, cluster = as.character(.data[[cluster_col]]), name = "ncells") %>%
    group_by(individual) %>%
    mutate(total_cells = sum(ncells)) %>%
    ungroup() %>%
    add_phase() %>%
    mutate(prop = ncells / total_cells)
}

prop_glm <- function(d) glm(cbind(ncells, total_cells - ncells) ~ Phase, family = binomial, data = d)

go_terms <- readRDS("Function Scripts/Dependencies/Term2gene_clown_go2.rds") %>%
  left_join(readRDS("Function Scripts/Dependencies/Term2name.rds"), by = "go_id")

# AUCell GO module score, averaged per individual
go_module <- function(term, o) {
  set.seed(1)
  genes <- intersect(go_terms$aocellaris_name[go_terms$go_id == term], rownames(o))
  message(length(genes), " genes found for: ", unique(go_terms$go_name[go_terms$go_id == term]))
  ranks <- AUCell_buildRankings(o@assays$RNA$data, plotStats = FALSE, verbose = FALSE)
  o$score <- as.numeric(getAUC(AUCell_calcAUC(setNames(list(genes), term), ranks, verbose = FALSE))[1, ])
  o@meta.data %>%
    group_by(individual, Status) %>%
    summarize(score = mean(score), .groups = "drop") %>%
    add_phase()
}

measures <- read.csv("Measures/2025_12_26 all_data.csv") %>%
  mutate(across(any_of(c("Behaviors_Day_2", "Time_Day_2", "Log_11KT", "Percent_Testicular",
                         "Percent_Ovarian", "Log10_Volume", "length_final_cm", "mass_final_cm",
                         "Change_Mass", "Change_Length", "Log10_Testicular_Estimate",
                         "Log10_Ovarian_Estimate")), as.numeric)) %>%
  add_phase(keep = all_phases)
```

# Fig. 1B-C (Table 1)

```{r fig1}
#### Fig. 1B-C (Table 1) ####
fig1_models <- list(
  `# parental acts` = lm(Behaviors_Day_2 ~ Phase, data = measures),
  `time in nest`    = lm(Time_Day_2 ~ Phase, data = measures),
  `log (11-KT)`     = lm(Log_11KT ~ Phase, data = measures),
  `% testicular`    = lm(Percent_Testicular ~ Phase, data = measures),
  `% ovarian`       = lm(Percent_Ovarian ~ Phase, data = measures),
  `volume`          = lm(Log10_Volume ~ Phase + length_final_cm, data = measures)
)
fig1_table <- stat_table(fig1_models, "Fig. 1B-C")
fig1_table

# Length-corrected volume (Fig. 1B bottom)
vol_len <- lm(Log10_Volume ~ length_final_cm, data = measures)
measures$resid_volume <- residuals(vol_len)[rownames(measures)]

ggplot(filter(measures, !is.na(resid_volume)), aes(Phase, resid_volume)) +
  geom_boxplot() +
  geom_point()

fig1_traits <- c("Behaviors_Day_2", "Time_Day_2", "Log_11KT",
                 "Percent_Testicular", "Percent_Ovarian", "resid_volume")
fig1_summary <- mean_se(measures, fig1_traits)
fig1_fold <- fold_table(measures, fig1_traits, no_ratio = "resid_volume")
fig1_summary
fig1_fold
```

# Fig. S1A-B

```{r figS1ab}
#### Fig. S1A-B ####
figS1ab_models <- list(
  `body mass`   = lm(mass_final_cm ~ Phase, data = measures),
  `body length` = lm(length_final_cm ~ Phase, data = measures)
)
figS1ab_table <- stat_table(figS1ab_models, "Fig. S1A-B")
figS1ab_summary <- mean_se(measures, c("mass_final_cm", "length_final_cm"))
figS1ab_fold <- fold_table(measures, c("mass_final_cm", "length_final_cm"))
figS1ab_table
figS1ab_summary
figS1ab_fold
```

# Fig. S1C-F

```{r figS1cf}
#### Fig. S1C-F ####
dat_experiment <- filter(measures, Condition == "Experiment")

figS1cf_models <- list(
  `change in mass`      = lm(Change_Mass ~ Phase, data = dat_experiment),
  `change in length`    = lm(Change_Length ~ Phase, data = dat_experiment),
  `log10 testis volume` = lm(Log10_Testicular_Estimate ~ Phase, data = measures),
  `log10 ovary volume`  = lm(Log10_Ovarian_Estimate ~ Phase, data = measures)
)
figS1cf_table <- stat_table(figS1cf_models, "Fig. S1C-F")
figS1cf_summary <- bind_rows(
  mean_se(dat_experiment, c("Change_Mass", "Change_Length")),
  mean_se(measures, c("Log10_Testicular_Estimate", "Log10_Ovarian_Estimate"))
)
figS1cf_fold <- bind_rows(
  fold_table(dat_experiment, c("Change_Mass", "Change_Length")),
  fold_table(measures, c("Log10_Testicular_Estimate", "Log10_Ovarian_Estimate"))
)
figS1cf_table
figS1cf_summary
figS1cf_fold
```

# snMultiome objects

```{r objects}
#### snMultiome objects ####
obj <- readRDS("~/Desktop/optimal_clustering_rna_only.rds")

# non-neuronal: 1 RGC, 2 OL, 11 MG, 13 OPC, 15 EC, 20 Leuko, 26 DG
neurons_only <- obj[, !obj$res0.8_50nn_40PC_45LSI %in% c(1, 2, 11, 13, 15, 20, 26)]

obj <- FindSubCluster(obj, 1, "harmony.wsnn", resolution = 0.2, subcluster.name = "sub_res0.2")
sub_1 <- subset(obj, final_clusters == 1)
Idents(sub_1) <- "sub_res0.2"
sub_1$cyto <- CytoTRACE(as.matrix(sub_1@assays$RNA$data))$CytoTRACE

sub_6 <- subset(obj, final_clusters == 6 & Status %in% c("M", "D", "F"))
sub_6$cyto <- CytoTRACE(as.matrix(sub_6@assays$RNA$data))$CytoTRACE
expr6 <- sub_6@assays$RNA$data
```

# Fig. 3A-B, S6A-B

```{r fig3ab}
#### Fig. 3A-B, S6A-B ####
counts_all <- cluster_counts(obj, "res0.8_50nn_40PC_45LSI")
counts_neuron <- cluster_counts(neurons_only, "res0.8_50nn_40PC_45LSI")

figS6a_table <- map(split(counts_all, counts_all$cluster), prop_glm) %>%
  stat_table("Fig. S6A") %>%
  mutate(q.value = p.adjust(`Type-III p-value`, "fdr"))
figS6b_table <- map(split(counts_neuron, counts_neuron$cluster), prop_glm) %>%
  stat_table("Fig. S6B") %>%
  mutate(q.value = p.adjust(`Type-III p-value`, "fdr"))

# Fig. 3A = 1_RGC among all cells; Fig. 3B = 6_POA_Mixed among neurons
fig3ab_table <- bind_rows(
  filter(figS6a_table, Trait == "1") %>% mutate(Figure = "Fig. 3A"),
  filter(figS6b_table, Trait == "6") %>% mutate(Figure = "Fig. 3B")
)

figS6a_summary <- map_dfr(split(counts_all, counts_all$cluster), mean_se, "prop", .id = "cluster")
figS6b_summary <- map_dfr(split(counts_neuron, counts_neuron$cluster), mean_se, "prop", .id = "cluster")
figS6a_fold <- map_dfr(split(counts_all, counts_all$cluster), fold_table, "prop", .id = "cluster")
figS6b_fold <- map_dfr(split(counts_neuron, counts_neuron$cluster), fold_table, "prop", .id = "cluster")

fig3ab_table
figS6a_table
figS6b_table
figS6a_summary
figS6b_summary
figS6a_fold
figS6b_fold
```

# Fig. 3C-D

```{r fig3cd}
#### Fig. 3C-D ####
chisq_enrichment <- function(d) {
  chi <- chisq.test(d$n)
  print(chi)
  d %>%
    mutate(expected = chi$expected,
           enrichment = n / expected,
           residual = (n - expected) / sqrt(expected),
           p_value = 2 * pnorm(-abs(residual)),
           p_adj = p.adjust(p_value, "BH"),
           signif = p_adj < 0.05 & enrichment > 1)
}

fig3c_deg <- read.csv("DEG Outputs/FINAL degs classified w singular.csv") %>%
  count(cluster) %>%
  chisq_enrichment()
fig3d_dar <- read.csv("Collaboration/all_clusters_DARs_peak_level_classified_with_support.csv") %>%
  count(cluster = cluster_id) %>%
  chisq_enrichment()
fig3c_deg
fig3d_dar
```

# Fig. 3E

```{r fig3e}
#### Fig. 3E ####
classifier_scores <- read.csv("Manuscript/Supplementary Tables/classifier_scores.csv") %>%
  filter(status == "D")

classifier_scores %>%
  group_by(cluster) %>%
  summarize(mean_score = mean(prediction), se_score = sd(prediction) / sqrt(n()))
classifier_scores %>%
  summarize(mean_score = mean(prediction), se_score = sd(prediction) / sqrt(n()))
```

# Fig. 4C-E

```{r fig4ce}
#### Fig. 4C-E ####
DimPlot(sub_1)

counts_rgc <- cluster_counts(sub_1, "sub_res0.2")

fig4de_models <- list(
  `1_NSC (1_1)` = prop_glm(filter(counts_rgc, cluster == "1_1")),
  `1_NP (1_0)`  = prop_glm(filter(counts_rgc, cluster == "1_0"))
)
fig4de_table <- stat_table(fig4de_models, "Fig. 4D-E")
fig4de_summary <- map_dfr(split(counts_rgc, counts_rgc$cluster), mean_se, "prop", .id = "cluster")
fig4de_fold <- map_dfr(split(counts_rgc, counts_rgc$cluster), fold_table, "prop", .id = "cluster") %>%
  filter(cluster %in% c("1_0", "1_1"))
fig4de_table
fig4de_summary
fig4de_fold
```

# Fig. 4F

```{r fig4f}
#### Fig. 4F ####
neuron_diff <- go_module("GO:0030182", subset(sub_1, sub_res0.2 == "1_0"))

fig4f_models <- list(`neuron differentiation (GO:0030182), 1_NP` = lm(score ~ Phase, data = neuron_diff))
fig4f_table <- stat_table(fig4f_models, "Fig. 4F")
fig4f_summary <- mean_se(neuron_diff, "score")
fig4f_fold <- fold_table(neuron_diff, "score")
fig4f_table
fig4f_summary
fig4f_fold
```

# Fig. S7B

```{r figS7b}
#### Fig. S7B ####
FeaturePlot(sub_1, "cyto")

cyto_sub1 <- sub_1@meta.data %>%
  filter(Status != "NRM") %>%
  group_by(individual, Status, sub_res0.2) %>%
  summarize(mean_cyto = mean(cyto), .groups = "drop") %>%
  mutate(sub_res0.2 = factor(sub_res0.2))

figS7b_models <- list(`CytoTRACE by 1_RGC subcluster` = lm(mean_cyto ~ sub_res0.2, data = cyto_sub1))
# previous model: lmer(mean_cyto ~ sub_res0.2 + (1 | individual), data = cyto_sub1)
figS7b_table <- stat_table(figS7b_models, "Fig. S7B", group = "sub_res0.2", contrasts = NULL)
figS7b_summary <- mean_se(cyto_sub1, "mean_cyto", group = "sub_res0.2")
figS7b_fold <- fold_table(cyto_sub1, "mean_cyto", group = "sub_res0.2")
figS7b_table
figS7b_summary
figS7b_fold
```

# Fig. 5B

```{r fig5b}
#### Fig. 5B ####
cyto_6 <- sub_6@meta.data %>%
  group_by(individual, Status) %>%
  summarize(cyto = mean(cyto), .groups = "drop") %>%
  add_phase()

fig5b_models <- list(`CytoTRACE, 6_POA_Mixed` = lm(cyto ~ Phase, data = cyto_6))
fig5b_table <- stat_table(fig5b_models, "Fig. 5B")
fig5b_summary <- mean_se(cyto_6, "cyto")
fig5b_fold <- fold_table(cyto_6, "cyto")
fig5b_table
fig5b_summary
fig5b_fold
```

# Fig. 5C-D

```{r fig5cd}
#### Fig. 5C-D ####
brain_dev <- go_module("GO:0007420", sub_6)
axon_guid <- go_module("GO:0008046", sub_6)
# axon_guid_alt <- go_module("GO:0097484", sub_6)  # GO term cited in the text

fig5cd_models <- list(
  `brain development (GO:0007420)` = lm(score ~ Phase, data = brain_dev),
  `axon guidance (GO:0008046)`     = lm(score ~ Phase, data = axon_guid)
  # , `axon guidance receptor activity (GO:0097484)` = lm(score ~ Phase, data = axon_guid_alt)
)
fig5cd_table <- stat_table(fig5cd_models, "Fig. 5C-D")
fig5cd_summary <- bind_rows(`GO:0007420` = mean_se(brain_dev, "score"),
                            `GO:0008046` = mean_se(axon_guid, "score"), .id = "go_term")
fig5cd_fold <- bind_rows(`GO:0007420` = fold_table(brain_dev, "score"),
                         `GO:0008046` = fold_table(axon_guid, "score"), .id = "go_term")
fig5cd_table
fig5cd_summary
fig5cd_fold
```

# Fig. 5G

```{r fig5g}
#### Fig. 5G ####
#### GO:0048167 regulation of synaptic plasticity, 6_POA_Mixed ####
syn_plas <- go_module("GO:0048167", sub_6)

ggplot(syn_plas, aes(Phase, score)) +
  geom_boxplot() +
  geom_point()

syn_plas_models <- list(`regulation of synaptic plasticity (GO:0048167)` = lm(score ~ Phase, data = syn_plas))
syn_plas_table <- stat_table(syn_plas_models, "6_POA_Mixed")
syn_plas_summary <- mean_se(syn_plas, "score")
syn_plas_fold <- fold_table(syn_plas, "score")
syn_plas_table
syn_plas_summary
syn_plas_fold
```

# Fig. 6A-C, S10A-E

```{r fig6ac}
#### Fig. 6A-C, S10A-E ####
genes_6 <- c(drd3 = "drd3", tacr3a = "tacr3a", cckb = "cckb", npy7r = "npy7r", nmbr = "nmbr",
             pgr = "pgr", gnrh1 = "LOC111571064", ar_like = "LOC111568069")

# Proportion of gene+ cells
gene_counts <- imap_dfr(genes_6, function(g, nm) {
  sub_6@meta.data %>%
    mutate(pos = expr6[g, ] > 0) %>%
    group_by(individual, Status) %>%
    summarize(n_pos = sum(pos), n_cells = n(), .groups = "drop") %>%
    mutate(gene = nm)
}) %>%
  add_phase() %>%
  mutate(prop = n_pos / n_cells)
gene_split <- split(gene_counts, factor(gene_counts$gene, levels = names(genes_6)))

fig6_prop_table <- map(gene_split, ~ glm(cbind(n_pos, n_cells - n_pos) ~ Phase,
                                         family = binomial, data = .x)) %>%
  stat_table("Fig. 6A-C, S10A-E")
fig6_prop_summary <- map_dfr(gene_split, mean_se, "prop", .id = "gene")
fig6_prop_fold <- map_dfr(gene_split, fold_table, "prop", .id = "gene")

# CytoTRACE of gene+ cells
cyto_genes <- c("drd3", "tacr3a", "cckb", "pgr", "ar_like")
cyto_pos <- map_dfr(cyto_genes, function(g) {
  sub_6@meta.data[expr6[genes_6[g], ] > 0, ] %>% mutate(gene = g)
}) %>%
  add_phase()
cyto_split <- split(cyto_pos, factor(cyto_pos$gene, levels = cyto_genes))

fig6_cyto_table <- map(cyto_split, ~ lmer(cyto ~ nCount_RNA + Phase + (1 | individual), data = .x)) %>%
  stat_table("Fig. 6A-C, S10A-E")
cyto_pos_ind <- cyto_pos %>%
  group_by(gene, individual, Phase) %>%
  summarize(cyto = mean(cyto), .groups = "drop")
fig6_cyto_summary <- map_dfr(split(cyto_pos_ind, cyto_pos_ind$gene), mean_se, "cyto", .id = "gene")
fig6_cyto_fold <- map_dfr(split(cyto_pos_ind, cyto_pos_ind$gene), fold_table, "cyto", .id = "gene")

fig6_prop_table
fig6_prop_summary
fig6_prop_fold
fig6_cyto_table
fig6_cyto_summary
fig6_cyto_fold
```

# Fig. 6D-F

```{r fig6df}
#### Fig. 6D-F ####
steroid <- read.csv("Manuscript/updatedcluster_6_steroid_receptor_SPECIFIC_PROMOTER_ZSCORES.csv") %>%
  group_by(individual, Phase = group) %>%
  summarize(mean_esr2b = mean(ESR2B_score), mean_ar = mean(AR_score),
            mean_pgr = mean(PGR_score), .groups = "drop") %>%
  filter(Phase %in% mif) %>%
  mutate(Phase = factor(Phase, levels = mif))

fig6df_models <- list(
  `ESR2B-associated expression` = lm(mean_esr2b ~ Phase, data = steroid),
  `AR-associated expression`    = lm(mean_ar ~ Phase, data = steroid),
  `PGR-associated expression`   = lm(mean_pgr ~ Phase, data = steroid)
)
fig6df_table <- stat_table(fig6df_models, "Fig. 6D-F")
fig6df_summary <- mean_se(steroid, c("mean_esr2b", "mean_ar", "mean_pgr"))
fig6df_fold <- fold_table(steroid, c("mean_esr2b", "mean_ar", "mean_pgr"),
                          no_ratio = c("mean_esr2b", "mean_ar", "mean_pgr"))
fig6df_table
fig6df_summary
fig6df_fold
```

# Named DEGs: inline statistics

```{r deg-inline}
#### Named DEGs: inline statistics ####
deg_genes <- list(
  `0` = c(igf2bp1 = "igf2bp1"),
  `1` = c(cyp19a1b = "LOC111577263"),
  # `1` = c(cyp19a1b = "cyp19a1b", gli1 = "gli1", uncx = "uncx", zic3 = "zic3", nfia = "nfia"),
  `6` = c(drd3 = "drd3", tacr3a = "tacr3a", npy7r = "npy7r", nmbr = "nmbr",
          cckb = "cckb", gnrh1 = "LOC111571064", pgr = "pgr")
)
cluster_names <- c(`0` = "0_Spall_Mixed", `1` = "1_RGC", `6` = "6_POA_Mixed")

deg_stats <- function(obj, cluster, genes, clustering = "res0.8_50nn_40PC_45LSI") {
  keep <- obj@meta.data[[clustering]] == cluster & obj$Status %in% c("M", "D", "F")
  counts <- obj@assays$RNA$counts[, keep]
  counts <- as.matrix(counts[Matrix::rowSums(counts) > 0, ])
  meta <- data.frame(subject = factor(obj$individual[keep]),
                     Status = factor(as.character(obj$Status[keep])))

  # size factors and dispersions from all genes in the cluster, as in the full DEG run
  sf <- scran::calculateSumFactors(counts, max.cluster.size = ncol(counts), positive = TRUE)
  disp <- glmGamPoi::glm_gp(counts, col_data = meta["Status"], size_factors = sf,
                            design = ~ Status, on_disk = FALSE)$overdispersion_shrinkage_list$ql_disp_estimate
  meta$log_sf <- log(sf)

  missing <- setdiff(genes, rownames(counts))
  if (length(missing)) warning("Cluster ", cluster, ": not found or all zero: ", paste(missing, collapse = ", "))
  genes <- genes[genes %in% rownames(counts)]

  lv <- levels(meta$Status)
  w <- function(a, b) as.numeric(lv == a) - as.numeric(lv == b)

  imap_dfr(genes, function(g, nm) {
    d <- mutate(meta, outcome = counts[g, ])
    m <- tryCatch(
      glmer(outcome ~ Status + (1 | subject), data = d, offset = log_sf,
            family = MASS::negative.binomial(theta = 1 / disp[match(g, rownames(counts))])),
      error = function(e) NULL)
    if (is.null(m)) return(tibble(cluster = cluster, Gene = nm, Symbol = g, error = TRUE))
    av <- Anova(m, type = 3)
    ct <- as.data.frame(contrast(emmeans(m, ~ Status),
                                 list(`I/M` = w("D", "M"), `F/M` = w("F", "M"), `F/I` = w("F", "D")),
                                 type = "response"))
    out <- tibble(cluster = cluster, Cluster = cluster_names[cluster], Gene = nm, Symbol = g,
                  Model = "glmer(counts ~ Phase + (1 | individual), NB, offset = log(size factor))",
                  `Wald Chisq` = av["Status", "Chisq"], df = av["Status", "Df"],
                  `Type-III p-value` = av["Status", "Pr(>Chisq)"])
    for (k in seq_len(nrow(ct))) {
      out[[paste(ct$contrast[k], "fold")]] <- ct$ratio[k]
      out[[paste(ct$contrast[k], "log2FC")]] <- log2(ct$ratio[k])
      out[[paste(ct$contrast[k], "p")]] <- ct$p.value[k]
    }
    out$converged <- is.null(m@optinfo$conv$lme4$messages)
    out$singular <- isSingular(m)
    out
  })
}

# q-values come from the full-cluster runs (FDR across all genes in the cluster)
saved_q <- map_dfr(names(deg_genes), function(cl) {
  read.csv(paste0("DEG Outputs/05_11_2025 Neg Bin w Doms New_clustering/cluster_", cl, ".csv")) %>%
    transmute(cluster = cl, Symbol = gene, saved_p = av_p.value,
              q.value = suppressWarnings(as.numeric(av_q.value)))
})

deg_table <- imap_dfr(deg_genes, ~ deg_stats(obj, .y, .x)) %>%
  left_join(saved_q, by = c("cluster", "Symbol")) %>%
  relocate(q.value, saved_p, .after = `Type-III p-value`)
deg_table
```

# Named DEGs: mean expression by phase

```{r deg-means}
#### Named DEGs: mean expression by phase ####
expr_by_ind <- imap_dfr(deg_genes, function(genes, cl) {
  keep <- obj$res0.8_50nn_40PC_45LSI == cl & obj$Status %in% c("M", "D", "F")
  genes <- genes[genes %in% rownames(obj)]
  ex <- obj@assays$RNA$data[genes, keep, drop = FALSE]
  imap_dfr(genes, function(g, nm) {
    tibble(individual = obj$individual[keep], Status = as.character(obj$Status[keep]), expr = ex[g, ]) %>%
      group_by(individual, Status) %>%
      summarize(expr = mean(expr), .groups = "drop") %>%
      add_phase() %>%
      mutate(Cluster = cluster_names[cl], Gene = nm)
  })
})

expr_summary <- expr_by_ind %>%
  group_by(Cluster, Gene, Phase) %>%
  summarize(mean = mean(expr), se = sd(expr) / sqrt(n()), n = n(), .groups = "drop")

expr_inline <- expr_summary %>%
  mutate(label = sprintf("%.3f ± %.3f", mean, se)) %>%
  dplyr::select(Cluster, Gene, Phase, label) %>%
  pivot_wider(names_from = Phase, values_from = label)

expr_fold <- expr_by_ind %>%
  group_by(Cluster, Gene) %>%
  group_modify(~ fold_table(.x, "expr")) %>%
  ungroup()

expr_summary
expr_inline
expr_fold
```

# gnrh1 (6_POA_Mixed)

```{r gnrh1}
#### gnrh1 (6_POA_Mixed): lm on per-individual mean expression ####
gnrh1_ind <- filter(expr_by_ind, Gene == "gnrh1")

gnrh1_models <- list(`gnrh1 expression, 6_POA_Mixed` = lm(expr ~ Phase, data = gnrh1_ind))
gnrh1_table <- stat_table(gnrh1_models, "Fig. S10C")
gnrh1_summary <- mean_se(gnrh1_ind, "expr")
gnrh1_fold <- fold_table(gnrh1_ind, "expr")
gnrh1_table
gnrh1_summary
gnrh1_fold
```

# Export (optional, not part of the analysis)

Writes every result table above to one Excel workbook (one sheet per table) next to this file, at full precision. If `writexl` isn't installed, it writes one CSV per table into a folder instead.

```{r export}
out_names <- ls(pattern = "(_table|_summary|_fold)$")
out_names <- c(out_names, intersect(c("fig3c_deg", "fig3d_dar", "expr_inline"), ls()))
out_list <- mget(out_names)
out_list <- out_list[vapply(out_list, is.data.frame, logical(1))]
# the two Fig. 3E summaries are printed above but not saved to objects; recompute them for export
if (exists("classifier_scores")) {
  out_list$fig3e_by_cluster <- classifier_scores %>% group_by(cluster) %>%
    summarize(mean_score = mean(prediction), se_score = sd(prediction) / sqrt(n()))
  out_list$fig3e_overall <- classifier_scores %>%
    summarize(mean_score = mean(prediction), se_score = sd(prediction) / sqrt(n()))
}
names(out_list) <- substr(names(out_list), 1, 31)

if (requireNamespace("writexl", quietly = TRUE)) {
  writexl::write_xlsx(out_list, "manuscript_statistics_outputs.xlsx")
  cat("Wrote manuscript_statistics_outputs.xlsx with", length(out_list), "sheets\n")
} else {
  dir.create("manuscript_statistics_outputs", showWarnings = FALSE)
  iwalk(out_list, ~ write.csv(.x, file.path("manuscript_statistics_outputs", paste0(.y, ".csv")),
                              row.names = FALSE))
  cat("Wrote", length(out_list), "CSVs to manuscript_statistics_outputs/\n")
}
```

# Session info

```{r session-info}
sessionInfo()
```
