GJG 06/06/2026 RGC Spatial locations

Here, I am going to see if our RGC subclusters correspond to spatially distinct 
populations of RGCs. One caveat here this will not be dissection aware while
I wait to hear back from Coltan re: which parts of sections should be excluded.

The steps will be
1) Subset visium object to be ONLY RGCs as defined by expression of markers
gfap, LOC111577263, sox2, or gja1b
1a) Run SCT on visium RGC object
2) Subset multiome object and subcluster
2a) Because ependymal cells are also gfap+ and crocc2, the main marker, does
not appear very strongly in the visium data, I will also include them
2b) 26, a dividing glial cluster will also be included
3) Use Seurat's label transfer functionality to go multiome_RGC -> Spatial data
4) Make a heatmap of regions 
5) Plot on spatial 

From prior knowledge, I hypothesize that 1_shha, 1_NSC, and 1_NP will be along 
the medial ventricles while 1_gfap will be more filamentous.

Libraries
```{r, echo = FALSE, include = FALSE}
library(Seurat)
library(ggplot2)
library(pheatmap)
library(tidyverse)
options(future.globals.maxSize = Inf)
library(glmGamPoi)
```

Read in visium data
```{r}
# ── 1. Load ──────────────────────────────────────────────────────────────────
vis <- readRDS('/Users/ggraham/Desktop/Visium/vis.rds')

# ── 2. Anatomical labels ──────────────────────────────────────────────────────
barcodes <- list.files(
  '/Users/ggraham/Desktop/multiome_poa/Visium/Groupings/Maruska Regions Data Informed/',
  full.names = TRUE
) %>%
  lapply(read.csv) %>%
  bind_rows()

ana <- setNames(barcodes$Anatomical, barcodes$Barcode)
seurat_barcodes <- gsub("_[0-9]+$", "", Cells(vis))
vis$anatomical <- unname(ana[seurat_barcodes])

# ── 4. Non-neuronal clusters (from marker genes) ─────────────────────────────
Idents(vis) <- 'seurat_clusters.projected0.4'
DotPlot(vis, features = c('elavl3','rbfox3a','gfap','LOC111577263','gja1b',
                          's100b','sox2',
                           'mpz','ptprc','p2ry12','cspg4','hbae5')) +
  coord_flip()

```

```{r, echo = FALSE, include = FALSE}
rgc_clusters = c(
  1,
  9,
  10,
  13,
  21,
  26,
  29
)

vis_rgc_test = subset(vis, seurat_clusters.projected0.4 %in% rgc_clusters)
marks = FindAllMarkers(vis_rgc_test)
```

```{r}
# Top 5 markers per cluster by p_val_adj
top5 <- marks %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  slice_min(order_by = p_val_adj, n = 5, with_ties = FALSE) %>%
  ungroup()

# DotPlot
DotPlot(
  vis_rgc_test,
  features = unique(top5$gene),
  group.by = "seurat_clusters.projected0.4"
) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
1 is certainly a radial glia 
9 I'm not sure none of its markers show up in the featureplot of the multiome object
10 is ependymal cells probably,
13 is like a new neuron or something
21 is a neuron, actually its vsst lol
26 looks like a rgc
29 is not

Evidence for above^^
```{r}
obj <- readRDS('~/Desktop/optimal_clustering_rna_only.rds')
```

```{r}
FeaturePlot(obj, c('apoa2',
                   'apoa1b',
                   'LOC111579859',
                   'LOC111579861',
                   'slc1a2b',
                   'ptprb',
                   'slc2a1b',
                   'prdx1',
                   'chga',
                   'cart3',
                   'sspo',
                   'tnfrsf19'), reduction = 'harmony_wnn.umap'
            ) & NoLegend() &
  theme(axis.title = element_blank(),
        axis.text = element_blank())
```
Subset visium object
```{r, echo = FALSE, include = FALSE}
vis_rgc = subset(vis,seurat_clusters.projected0.4 %in% c(1, 10,26) )
rm(vis_rgc_test)
rm(vis)
```

Run SCT, save
```{r, echo = FALSE, include = FALSE}
vis_rgc= SCTransform(vis_rgc, assay = 'Spatial.Polygons',
                          conserve.memory = T,
                          vst.flavor = "v2",
                          ncells = 5000)
  
#saveRDS(vis_rgc,'/Users/ggraham/Desktop/Visium/vis_rgc.rds' )

```

Subset and subcluster ependymoglia
```{r, echo = FALSE}
obj = FindSubCluster(obj,
                     1, 
                     'harmony.wsnn',
                     resolution = 0.2,
                     subcluster.name = 'sub_res0.2')

obj_rgc = subset(obj, final_clusters %in% c(1,
                                            15,
                                            26))

obj_rgc$sub_res0.2 <- recode(obj_rgc$sub_res0.2,
  "1_0" = "1_NP",
  "1_1" = "1_NSC",
  "1_2" = "1_shha",
  "1_3" = "1_gfap",
  "15" = 'Ependymal',
  '26' = 'Dividing'
)
Idents(obj_rgc) = 'sub_res0.2'
rm(obj)
DimPlot(obj_rgc, label = T)
```
Perfect, now, label transfer time

Balance object so that larger populations do not overwhelm the transfer
```{r}
cells_per_cluster <- 200

set.seed(42)
cells_keep <- obj_rgc@meta.data %>%
  tibble::rownames_to_column("cell") %>%
  group_by(sub_res0.2) %>%
  group_modify(~ slice_sample(.x, n = min(cells_per_cluster, nrow(.x)))) %>%
  pull(cell)


obj_rgc_balanced <- subset(obj_rgc, cells = cells_keep)

```


```{r, echo = FALSE, include = FALSE}
# Find Transfer anchors using SCT assay and transfer data
DefaultAssay(obj_rgc_balanced)        <- "SCT"
Idents(obj_rgc_balanced) = 'sub_res0.2'
# now run transfer on balanced reference
anchors <- FindTransferAnchors(
  reference            = obj_rgc_balanced,
  query                = vis_rgc,
  normalization.method = "SCT",
  reference.reduction  = "pca",
  reduction            = "pcaproject",
  dims                 = 1:30,
    query.assay          = "SCT"      

)


vis_rgc <- TransferData(
  anchorset        = anchors,
  query            = vis_rgc,
  reference        = obj_rgc_balanced,
  refdata          = list(transferred_cluster = "sub_res0.2"),
  weight.reduction = "pcaproject",
  dims             = 1:30
)

```

```{r}
vis_rgc$seurat_clusters.projected0.4 = droplevels(vis_rgc$seurat_clusters.projected0.4 )
# score distribution split by whether assignment was high confidence
vis_rgc$transfer_confident <- 
  vis_rgc$predicted.transferred_cluster.score > 0.5

cat("\nHigh confidence assignments:", 
    sum(vis_rgc$transfer_confident), "\n")
cat("Low confidence assignments:", 
    sum(!vis_rgc$transfer_confident), "\n")

```
Very good confidence

Plot against visium clusters
```{r}
  vis_rgc$predicted.cluster=   vis_rgc$predicted.transferred_cluster[vis_rgc$transfer_confident==T]

agreement = table(vis_rgc$predicted.cluster,
                  vis_rgc$seurat_clusters.projected0.4)

pheatmap(
  agreement / rowSums(agreement),               # each region sums to 1
  cluster_rows = F,
  cluster_cols = F,
  border_color = NA,
  color        = colorRampPalette(c("grey95", "orange", "darkred"))(100)
)

```
Ok so yeah they are overwhelmingly aligned to cluster 1 not a surprise, it looks
like ependymal cells (15) are also a part of that cluster, dividing glia are kind of evenly divided.

Is there a spatial difference

```{r}
renamed = c(
  'could be nPPa'= 'What',
  'could be nMMp' = 'nMMp',
  'could be nPMP' = 'nPMp',
  'dorsolateral nucleus' = 'Dln-r',
  'dorsolateral nucleus caudal part' = 'Dln-c',
  'Dl-v2 dorsal part' = 'Dl-v2d',
  'TGn' = 'TGN'
)

renaming_col = ifelse(vis_rgc$anatomical %in% c(names(renamed)), renamed[vis_rgc$anatomical], vis_rgc$anatomical)  
vis_rgc$anatomical_renamed = renaming_col

  vis_rgc$predicted.cluster=   vis_rgc$predicted.transferred_cluster[vis_rgc$transfer_confident==T]

agreement = table(vis_rgc$predicted.cluster,
                  vis_rgc$anatomical_renamed)

pheatmap(
  t(agreement / rowSums(agreement)),               # each region sums to 1
  cluster_rows = F,
  cluster_cols = F,
  border_color = NA,
  color        = colorRampPalette(c("grey95", "orange", "darkred"))(100),
  fontsize=4)


```
Visually, a lot of confidence for the rgcs in OT which isnt surprising because
there is so many of them there that I guess it is much more confident. In POA,
seems to be more ependymal cells and 1_2 which is 1_shha, lets just plot it though

```{r}
SpatialDimPlot(vis_rgc, group.by = 'predicted.cluster',
               images = 's_6P17.polygons',
               pt.size.factor = 0.7)
```

```{r}
SpatialDimPlot(vis_rgc, group.by = 'predicted.cluster',
               images = 's_4P10.polygons',
               pt.size.factor = 0.7)
```
1_1 seems to be more posterior

```{r}
SpatialDimPlot(vis_rgc, group.by = 'predicted.cluster',
               images = 's_3P5.polygons',               
               pt.size.factor = 0.7)
```

```{r}
SpatialDimPlot(vis_rgc, group.by = 'predicted.cluster',
               images = 's_4P5.polygons',
                              pt.size.factor = 0.7)
```



