To obtain pseudobulked data, gene counts were summed across all cells belonging to the same cell cluster and sample resulting in a matrix of 30,704 genes by 24 samples by 53 cell clusters. Differential expression analysis was then performed with edgeR

```{r message=  FALSE}
library(Seurat)
library(tidyverse)
library(edgeR)
```

read in object

```{r}
obj <- readRDS('~/Desktop/snRNA-seq R Files 122524/RNA Object.rds')
```

subset obj

```{r}
subset_obj <- subset(obj, Status %in% c('M','D','F'))
subset_obj <- subset(subset_obj, harmony.wnn_res0.4_clusters != 15 & harmony.wnn_res0.4_clusters != 30)

```

```{r}
DimPlot(subset_obj, reduction = 'harmony_wnn.umap', label = T)
```

So I think I want to do is: for each cluster, aggregate counts by individual, and include a column for sex and pool

```{r}
# Make a shorter cluster name so my life is easier
subset_obj$cluster = subset_obj$harmony.wnn_res0.4_clusters
# Initialize a list to store aggregated counts in 
counts_cluster <- list()
for(i in unique(subset_obj$cluster)){ # For all clusters

  # Make a list that stores the indices of which cells belong to that cluster
  cluster_cells = which(subset_obj$cluster == i)
  
  # Make a counts matrix for all genes in that cluster
  counts <- subset_obj@assays$RNA$counts[, cluster_cells] 
  
  # Extract meta data corresponding to those cells
  individual_ids <- subset_obj$individual[cluster_cells]
  Status_ids <- subset_obj$Status[cluster_cells]
  orig.ident_ids <- subset_obj$orig.ident[cluster_cells]
  
  # Initialize a genes list
  genes <- rownames(counts)
  
  # Get unique individuals
  unique_individuals <- unique(individual_ids)
  
  # Make an empty matrix where rows are genes and columns are UNIQUE individuals
  agg_counts <- matrix(0, nrow=length(genes), ncol=length(unique_individuals))
  
  # Give row and column names
  rownames(agg_counts) <- genes
  colnames(agg_counts) <- unique_individuals
  
  # Create metadata dataframe with the same order as unique_individuals
  meta_data <- data.frame(
    individual = unique_individuals,
    Status = character(length(unique_individuals)),
    orig.ident = character(length(unique_individuals)),
    stringsAsFactors = FALSE
  )
  
  # Fill in metadata in the same order as columns
  for(idx in 1:length(unique_individuals)){
    j <- unique_individuals[idx]
    # Find the indices that correspond to this specific individual
    individual_cells <- which(individual_ids == j)
    
    # Sum the counts for each gene across all cells from this individual
    agg_counts[, j] <- rowSums(counts[, individual_cells, drop=FALSE])
    
    # Store metadata for this individual (using first occurrence)
    first_cell <- individual_cells[1]
    meta_data$Status[idx] <- Status_ids[first_cell]
    meta_data$orig.ident[idx] <- orig.ident_ids[first_cell]
  }
  
  # Verify that the order matches
  stopifnot(all(colnames(agg_counts) == meta_data$individual))
  
  # Store the aggregated counts and metadata
  counts_cluster[[i]] <- list(
    aggregated_counts = t(agg_counts), #transpose for later usability
    meta.data = meta_data
  )
}
```

```{r}

```
