---
title: 'Using EmptyDropsMultiome called barcodes in souporcell, then cross-referencing with demuxlet sample calls'
author: Jenny Drnevich and Jessica Holmes, HPCBio, Roy J. Carver Biotechnology Center, University of Illinois,
  Urbana-Champaign
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    self_contained: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

```

## Load packages

```{r libs, message = FALSE, warning = FALSE}
if (!require("dplyr")) {
    install.packages("dplyr")
    library(dplyr)
}
```


## Read in demuxlet and souporcell for pool 1


```{r echo=TRUE, message=FALSE, warning=FALSE}

s1.d <- read.csv("demuxlet/Pool_1/s1_gex.best", sep ="\t", header = T)
s1.s <- read.csv("EmptyDropsMultiome/s1/clusters.tsv", sep ="\t", header = T)

# First lowercase the demuxlet columns
colnames(s1.d) <- tolower(colnames(s1.d))

# Add prefixes for each column, except for barcode
colnames(s1.d)[-2] <- paste("dm",colnames(s1.d)[-2], sep = "_")
colnames(s1.s)[-1] <- paste("sc",colnames(s1.s)[-1], sep = "_")

# Join by barcodes
s1.m <- left_join(s1.d, s1.s)

# check overlaps of singlet/doublet calls

table(s1.m$dm_droplet.type)
table(s1.m$sc_status)

table(s1.m$dm_droplet.type, s1.m$sc_status)

```


## Cut down to only souporcell singlets

We are going to use souporcell's singlet calls only, so remove all other cell barcodes

```{r echo=TRUE, message=FALSE, warning=FALSE}

s1.m <- s1.m[s1.m$sc_status == "singlet",]

temp1 <- table(s1.m$dm_sng.best.guess, s1.m$sc_assignment)
temp1


```


Extremely good overlap between souporcell's assignment to clusters 0-3 and demuxlet's sample assignment! Use this information to assign sample IDs to each barcode (note: souporcell's cluster assignment overrides demuxlet's )

```{r echo=TRUE, message=FALSE, warning=FALSE}

temp2 <- as.data.frame(temp1)
temp2 <- slice_max(temp2, order_by =  Freq, n = 4, with_ties = FALSE) %>% arrange(Var2)
temp <- s1.m[,c("barcode", "sc_assignment")]
names(temp)[2] <- "Sample"
temp$Sample <- factor(temp$Sample)
levels(temp$Sample) <- as.character(temp2$Var1)

write.csv(temp, "EmptyDropsMultiome/s1_barcodes_sampleIDs.csv")

table(temp$Sample)

```

## Repeat for pools 2, 3, 5 & 6

Pool 4 is the one where one of the 4 individuals did not get DNA sequenced, so demuxlet did not have it's variants to be able to assign cells to it. The other pools can follow the same codes as pool 1

```{r echo=TRUE, message=FALSE, warning=FALSE}

for (i in c(2,3,5,6)){
  dm <- read.csv(paste0("demuxlet/Pool_", i,"/s",i, "_gex.best"), sep ="\t", header = T)
  sc <- read.csv(paste("EmptyDropsMultiome/s", i,"/clusters.tsv", sep = ""), sep ="\t", header = T)
  colnames(dm) <- tolower(colnames(dm))
  colnames(dm)[-2] <- paste("dm",colnames(dm)[-2], sep = "_")
  colnames(sc)[-1] <- paste("sc",colnames(sc)[-1], sep = "_")
  dm.sc <- left_join(dm, sc)

  sx.m <- dm.sc[dm.sc$sc_status == "singlet",]
  temp1 <- table(sx.m$dm_sng.best.guess, sx.m$sc_assignment)
  print(paste("pool", i))
  print(temp1)
  temp2 <- as.data.frame(temp1)
  temp2 <- slice_max(temp2, order_by =  Freq, n = 4, with_ties = FALSE) %>% arrange(Var2)
  temp <- sx.m[,c("barcode", "sc_assignment")]
  names(temp)[2] <- "Sample"
  temp$Sample <- factor(temp$Sample)
  levels(temp$Sample) <- as.character(temp2$Var1)
  write.csv(temp, paste0("EmptyDropsMultiome/s",i,"_barcodes_sampleIDs.csv"))
  
  print(paste("pool", i))
  print(table(temp$Sample))
  }

```

## Modify for pool 4

Pool 4 contains B21F, T13D, T5D and GH. GH was the one that did not get it's DNA sequenced. Demuxlet had a much harder time with the pool and called a much higher percentage of doublets than the other pools. 


```{r echo=TRUE, message=FALSE, warning=FALSE}

s4.d <- read.csv("demuxlet/Pool_4/s4_gex.best", sep ="\t", header = T)
s4.s <- read.csv("EmptyDropsMultiome/s4/clusters.tsv", sep ="\t", header = T)

# First lowercase the demuxlet columns
colnames(s4.d) <- tolower(colnames(s4.d))

# Add prefixes for each column, except for barcode
colnames(s4.d)[-2] <- paste("dm",colnames(s4.d)[-2], sep = "_")
colnames(s4.s)[-1] <- paste("sc",colnames(s4.s)[-1], sep = "_")

s4.m <- left_join(s4.d, s4.s)

# check overlaps

table(s4.m$dm_droplet.type)
table(s4.m$sc_status)

table(s4.m$dm_droplet.type, s4.m$sc_status)


table(s4.m$dm_droplet.type, s4.m$dm_sng.best.guess)

# All have many doublets that are likely the 4th sample

# Cut down to only souporcell singlets
s4.m <- s4.m[s4.m$sc_status == "singlet",]

temp1 <- table(s4.m$dm_sng.best.guess, s4.m$sc_assignment)
temp1


```


Looks like souporcell's cluster 3 is the missing GH sample. What does this look like if we only use demuxlet's singlet calls as well?

```{r echo=TRUE, message=FALSE, warning=FALSE}

s4.m2 <- s4.m[s4.m$dm_droplet.type == "SNG",]
table(s4.m2$dm_sng.best.guess, s4.m2$sc_assignment)

```


I feel very comfortable calling cluster 3 as the GH sample. 


```{r echo=TRUE, message=FALSE, warning=FALSE}

temp2 <- as.data.frame(temp1)
temp2 <- slice_max(temp2, order_by =  Freq, n = 4, with_ties = FALSE) %>% arrange(Var2)
temp2

temp2$Var1 <- as.character(temp2$Var1)
temp2$Var1[4] <- "GH"

temp <- s4.m[,c("barcode", "sc_assignment")]
names(temp)[2] <- "Sample"
temp$Sample <- factor(temp$Sample)
levels(temp$Sample) <- as.character(temp2$Var1)

write.csv(temp, "EmptyDropsMultiome/s4_barcodes_sampleIDs.csv")

table(temp$Sample)

```

## Final tallies

```{r echo=TRUE, message=FALSE, warning=FALSE}
i <- 1
allcells <- read.csv(paste0("EmptyDropsMultiome/s",i,"_barcodes_sampleIDs.csv"))
allcells$pool <- i

for(i in 2:6){
  temp <- read.csv(paste0("EmptyDropsMultiome/s",i,"_barcodes_sampleIDs.csv"))
  temp$pool <- i
  allcells <- rbind(allcells, temp)
}

# Total number of cells
nrow(allcells)

# Number of cells per pool

table(allcells$pool)

# Number of cells per sample

table(allcells$Sample)

```

