---
title: "subclustering_cluster_2_03_20_2025"
format: html
editor: visual
bibliography: references.bib
---

```         
quarto::quarto_render(here::here("C:/Users/Gabe/Desktop/multiome_poa/Subclustering/Cluster 2/radial_glia_WGCNA_analysis_03_20_2025.qmd"),                       execute_dir = here::here(),                       output_format = "all", as_job = TRUE)
```

Libraries

```{r}
library(here)
#| echo: false
setwd('C:/Users/Gabe/Desktop/multiome_poa')

{
  library(parallel)
  library(clusterProfiler)
  library(blme)
  library(Seurat)
  library(tidyverse)
  library(tidyr)
  library(lme4)
  library(dplyr)
  library(MASS)
  library(SeuratObject)
  library(Signac)
  library(glmGamPoi)
  library(scran)
  library(parallel)
  library(factoextra)
  library(readxl)
  library(factoextra)
  library(forcats)
  library(ggrepel)
  library(biomaRt)
  library(openxlsx)
  library(emmeans)
  library(CytoTRACE)
  library(ggrepel)
  library(tidyverse)
  library(cowplot)
  library(patchwork)
  library(Polychrome)
  library(scCustomize)

P40 <- createPalette(40, c("#FF0000", "#00FF00", "#0000FF"), range = c(30, 80))
swatch(P40)
names(P40) <- NULL

mean_expression_cluster_plot<- readRDS('Functions/mean_expression_cluster_plot.rds')
prop_cluster_plot<- readRDS( 'Functions/prop_cluster_plot.rds')
define_degs_prop<- readRDS('Functions/define_degs_prop.rds')
mean_expression_cluster_data<- readRDS('Functions/mean_expression_cluster_data.rds')
prop_deg_function.rds<- readRDS('Functions/DEG_functions/prop_deg_function.rds')
define_behavior_degs<- readRDS('Functions/define_behavior_degs')
clown_go<- readRDS('Functions/clown_go')
define_degs<- readRDS('Functions/define_degs')

}

```

### Load in the Seurat Object and Subset to Radial Glia, find subclusters

```{r}
obj <- readRDS("C:/Users/Gabe/Desktop/RNA Object.rds")
radial_glia <- subset(obj, harmony.wnn_res0.4_clusters==2)
radial_glia <- FindSubCluster(radial_glia, cluster = 2, subcluster.name = 'sub', graph.name = 'harmony.wsnn')

```

Read in Markers from Previous Analysis

```{r}
glia_markers <- readRDS('Subclustering/Cluster 2/glia subcluster markers.rds')

```

```{r}
zack_markers <- read.csv('Subclustering/Cluster 2/zack_glia_markers_ocellaris.csv')

```

### **Visualize Zack's Markers**

Quiescent

```{r}
Idents(radial_glia)<- 'sub'
DotPlot(radial_glia, unique(zack_markers$gene[zack_markers$rg_state=='quiescent']))+
  coord_flip()
```

Cycling

```{r}
DotPlot(radial_glia, unique(zack_markers$gene[zack_markers$rg_state=='cycling']))+
  coord_flip()

```

Neuronal Differentiation

```{r}
DotPlot(radial_glia, unique(zack_markers$gene[zack_markers$rg_state=='neuronal differentiation']))+
  coord_flip()

```

Ok to me, classifications are as follows so far:

2_0: Quiescent

2_1: Cycling/Quiescent

2_2: Quiescent

2_3: Strongly Quiescent

2_4: Cycling

2_5: Cycling / Quiescent

2_6: Quiescent

2_7: Quiescent / Neuronal Differentiation

2_8: Quiescent

2_9: Neuronal Differentiation

### **What does WGCNA have to say about this?**

Set Up for WGCNA

```{r}
library(hdWGCNA)
radial_glia <- SetupForWGCNA(
  radial_glia,
  gene_select = "fraction",
  fraction = 0.05, 
  wgcna_name = "radial_glia" 
)

```

Construct metacells

```{r}
radial_glia <- MetacellsByGroups(
  seurat_obj = radial_glia,
  group.by = c("sub"),
  reduction = 'harmony_wnn.umap', 
  k = 10, #reduce k  due to small number of cells
  max_shared = 5, 
  ident.group = 'sub',
  min_cells = 25 #set to 25 such that all subclusters can be included
)

```

Normalize metacell expression

```{r}
radial_glia <- NormalizeMetacells(radial_glia)

```

Coexpression network analysis

```{r}
clusters_list <- unique(radial_glia$sub
                       )
radial_glia <- SetDatExpr(
  radial_glia,
  group_name = paste0(clusters_list),
  #changing this to be all clusters
  group.by='sub', 
  assay = 'RNA', 
  layer = 'data' 
)
```

Select soft power threshold

```{r}
radial_glia <- TestSoftPowers(
  radial_glia,
  networkType = 'signed' #not sure if this should be altered either
)
```

Plot soft power results

```{r}
plot_list <- PlotSoftPowers(radial_glia)
wrap_plots(plot_list, ncol=2)
```

5 is our soft power threshold

```{r}
radial_glia <- ConstructNetwork(
  radial_glia,
  tom_name = 'radial_glia_subclusters_03_20_2025_TOM',
tom_outdir = 'A:/anemonefish_multiome_radial_glia_03_20_2025/',
overwrite_tom = T,
soft_power = 5
  )

```

Plot

```{r}
PlotDendrogram(radial_glia, main='RGC hdWGCNA Dendrogram')

```

3 main modules interesting

Module Eigengenes

```{r}
radial_glia <- ModuleEigengenes(
 radial_glia,
 group.by.vars="sub"
)
```

```{r}
#| echo: false

radial_glia <- readRDS(file='A:/anemonefish_multiome_radial_glia_03_20_2025/radial_glia_object.rds')
library(hdWGCNA)
```

Harmonized and non-harmonized module eigengenes

```{r}
# harmonized module eigengenes:
hMEs <- GetMEs(radial_glia)
# module eigengenes:
MEs <- GetMEs(radial_glia, harmonized=FALSE)

```

Compute eigengene based connectivity

```{r}
clusters_list <- unique(radial_glia$sub
                       )
radial_glia@misc[["radial_glia"]][["wgcna_net"]][["TOMFiles"]] <-"A:/anemonefish_multiome_radial_glia_03_20_2025/radial_glia_subclusters_03_20_2025_TOM_TOM.rda"

radial_glia <- ModuleConnectivity(
  radial_glia,
  group.by = 'sub', group_name =  paste0(clusters_list)
  
)

```

Get module assignment table

```{r}
modules <- GetModules(radial_glia) %>% subset(module != 'grey')
```

Get hub genes

```{r}
hub_df <- GetHubGenes(radial_glia, n_hubs = 10)

```

Compute hub signature score

```{r}
radial_glia <- ModuleExprScore(
  radial_glia,
  n_genes = 25,
  method='UCell'
)

```

Plot hMEs

```{r}
plot_list <- ModuleFeaturePlot(
  radial_glia,
  reduction = 'harmony_wnn.umap',
  features='hMEs', # plot the hMEs
  order=TRUE 
)
wrap_plots(plot_list, ncol=3)
```

```{r}
rgc_modules = ModuleRadarPlot(
  radial_glia,
  barcodes =  radial_glia@meta.data %>% rownames(),
  axis.label.size=4,
  grid.label.size=3
)
rgc_modules
```

```{r}
#setwd('C:/Users/Gabe/Desktop/multiome_poa')
#ggsave(plot = rgc_modules,
#       file = "rgc_modules.svg",
#       device = "svg",
#       units = "in",
#       width = 8,
#       height = 6,
#       path = "Bachelors Thesis/Plots/Figure 5")

```

Fascinating, blue module corresponds strongly to 2_9 our newborn neuron cluster, interestingly, brown module corresponds with 2_9 and 2_7 suggesting these may be progenitors or something. Turquoise and yellow I am less clear on.

Analyze what the modules are enriched for

```{r}
modules <- GetModules(radial_glia)
mods <- levels(modules$module); mods <- mods[mods != 'grey']

radial_glia@meta.data <- cbind(radial_glia@meta.data, MEs)

```

```{r}
DotPlot(radial_glia, features=mods, group.by = 'sub')+
  coord_flip()
```

Ok well 2_9 is enriched for everything which is not surprising as it is some kind of stem cell, but the blue and brown modules are the most specific. I think this somewhat aligns with some of my cells being in between quiscent and cycling in my above classification

Analyze them

```{r}
clown_go <- readRDS("C:/Users/Gabe/Desktop/multiome_poa/Functions/clown_go") 
module_go_function <- function(module_name,...){
  genes <-  modules$gene_name[modules$module==module_name]
 return(clown_go(genes,...)%>%dotplot())
  }

```

```{r}
setwd('C:/Users/Gabe/Desktop/multiome_poa')
module_go_function('brown', whole.list = modules$gene_name)

```

```{r}
setwd('C:/Users/Gabe/Desktop/multiome_poa')
module_go_function('blue')
```

```{r}
setwd('C:/Users/Gabe/Desktop/multiome_poa')
module_go_function('yellow')
```

Ok this is really unexpected, yellow??

Looking at the previous graphs this might make sense, 2_9 is enriched in this one too, also 2_2 and 2_3, very interesting because my marker genes suggest 2_2 and 3 are quiescent. I wonder if maybe they are progenitors for different lineages?

```{r}
setwd('C:/Users/Gabe/Desktop/multiome_poa')
module_go_function('turquoise')
```

Plot number of genes per module

```{r}
modules%>%
  group_by(module)%>%
  subset(module!='grey')%>%
  summarize(n = n())%>%
  plot()
```

Ok this is super interesting. Blue is enriched for 2_9 because 2_9 expresses everything, as for turquoise I have no idea. It's crazy how small yellow is and how strong its GO signal is

Plot network

```{r}
ModuleNetworkPlot(
  radial_glia,
  outdir = 'A:/',
   n_inner = 20, # number of genes in inner ring
    n_outer = 30, # number of genes in outer ring
    n_conns = Inf, # show all of the connections
    plot_size=c(10,10), # larger plotting area
    vertex.label.cex=1 # font size
)
```

```{r}
options('future.globals.maxSize'=Inf)
HubGeneNetworkPlot(
  radial_glia,
  n_hubs = 3, n_other=5,
  edge_prop = 0.75,
  mods = 'all'
)

```

```{r}
radial_glia <- RunModuleUMAP(
  radial_glia,
  n_hubs = length(unique(modules$module))-1, 
  n_neighbors=15, 
  min_dist=0.1 
)
```

```{r}
umap_df <- GetModuleUMAP(radial_glia)

ggplot(umap_df, aes(x=UMAP1, y=UMAP2)) +
  geom_point(
   color=umap_df$color, # color each point by WGCNA module
   size=umap_df$kME*2 # size of each point based on intramodular connectivity
  ) +
  umap_theme()

```

Gorg

```{r}
ModuleUMAPPlot(
  radial_glia,
  edge.alpha=0.25,
  sample_edges=TRUE,
  edge_prop=0.1, 
  label_hubs=2 ,
  keep_grey_edges=FALSE,
    return_graph = F
)
```

```{r}
#saveRDS(radial_glia, file='A:/anemonefish_multiome_radial_glia_03_20_2025/radial_glia_object.rds')
```

### ---

Also I just discovered this function of quarto

@oldham2006 which links to my zotero somehow so cool
