```{r}
library(Seurat)
library(CytoTRACE)
library(lme4)
library(tidyverse)
```

```{r}
obj = readRDS("~/Desktop/optimal_clustering_rna_only.rds")
```

here, I want to compare two methods of subclustering cluster 0, one is using FindSubClusters and the other is subsetting, finding new top variable genes etc and then using findclusters

```{r}
findsub_0 = FindSubCluster(obj, 0, graph.name = 'harmony.wsnn')%>%
  subset(final_clusters ==0)
```

```{r}
DimPlot(findsub_0, reduction = 'harmony_wnn.umap', group = 'sub.cluster')
```
A hunch I've had for a while now is one of these clusters is very very immature, so lets see

```{r}
findsub_0$cyto = CytoTRACE(findsub_0@assays$RNA$data%>%as.matrix())$CytoTRACE
```
```{r}
findsub_0@meta.data$Status = factor(findsub_0@meta.data$Status, levels = c('NRM','M','D','E','NF','F'))
dat = findsub_0@meta.data%>%
  group_by(individual, Status, sub.cluster)%>%
  summarize(mean_cyto = mean(cyto),
            se_cyto =sd(cyto)/sqrt(n()))
```
```{r}
ggplot(dat, aes(x = sub.cluster, y = mean_cyto, color = Status))+
  geom_boxplot()
```
So yeah I was basically dead wrong

Proportion differences?
```{r}
allcell_findsub_0 = findsub_0@meta.data%>%
  group_by(individual, Status)%>%
  summarize(total_cells = n())

subcell_findsub_0 = findsub_0@meta.data%>%
  group_by(individual, Status, sub.cluster)%>%
  summarize(cells = n())

prop_cells_findsub0 = subcell_findsub_0%>%
  right_join(allcell_findsub_0, by = 'individual')

prop_cells_findsub0%>%glimpse()

```
```{r}
ggplot(prop_cells_findsub0, aes(x = sub.cluster, y = cells/total_cells, color = Status.x))+
  geom_boxplot()
```

I think the only interesting things here are 0_1 and 0_3, 0_3 also has an interesting CytoTRACE profile

Lets look at markers then
```{r}
findsub_0_markers = FindAllMarkers(findsub_0, group.by = 'sub.cluster', only.pos = T)
```

```{r}
top10_markers <- findsub_0_markers %>%
  group_by(cluster) %>%
  slice_min(order_by = p_val_adj, n = 3)

DotPlot(findsub_0, features = unique(top10_markers$gene), group.by = 'sub.cluster') + 
  coord_flip()
```
0_0 and 0_2 are GABAergic (slc32a1), 0_1 and 0_3 are Glutamatergic

I guess the question now is why are these cells clustering together if some are glutamatergic and others are gabaergic

```{r}
DimPlot(obj)
```
```{r}
DimPlot(findsub_0, group.by ='sub.cluster')
```

```{r}
very_strong_markers_findsub_0 = subset(findsub_0_markers, pct.1>0.9)
```
Still not clear, seems like there is evidence of migration but I'm not convinced

Now to try with reclustering
```{r}
findclust_0 = subset(obj, final_clusters ==0)
```

```{r}
findclust_0 <- FindVariableFeatures(findclust_0, selection.method = "vst", nfeatures = 5000)
findclust_0 <- ScaleData(findclust_0)
findclust_0 <- RunPCA(findclust_0,features = VariableFeatures(object = findclust_0))
```

```{r}
ElbowPlot(findclust_0)
```

I guess 8? Strong arguments also for 2

```{r}
DimHeatmap(findclust_0, dims = 1:10, cells = 500, balanced = TRUE)

```
I mean there is still strong signal at 9 so I guess lets go with 8

```{r}
findclust_0 <- FindNeighbors(findclust_0, dims = 1:8)

```

```{r}
set.seed(0)
findclust_0 <- FindClusters(findclust_0, resolution = 0.5)
```

```{r}
set.seed(0)
findclust_0 <- RunUMAP(findclust_0, dims = 1:8)
```
```{r}
DimPlot(findclust_0,reduction="umap",group.by="seurat_clusters",label=T)
```
This looks overclustered to me

```{r}
set.seed(0)
findclust_0 <- FindClusters(findclust_0, resolution = 0.1)
findclust_0 <- RunUMAP(findclust_0, dims = 1:8, reduction.name = 'res.1')
DimPlot(findclust_0,reduction="umap",group.by="RNA_snn_res.0.1",label=T)
```
It seems like these clusters are robust so now lets analyze them
```{r}
Idents(findclust_0) = 'RNA_snn_res.0.1'
findclust_0_markers =FindAllMarkers(findclust_0, only.pos = T)
```
```{r}
top10_markers_2 <- findclust_0_markers %>%
  group_by(cluster) %>%
  slice_min(order_by = p_val_adj, n = 3, with_ties = F)

DotPlot(findclust_0, features = unique(top10_markers_2$gene), group.by = 'RNA_snn_res.0.1') + 
  coord_flip()
```
```{r}
findclust_0$cyto = CytoTRACE(findclust_0@assays$RNA$data%>%as.matrix())$CytoTRACE

```
```{r}
findclust_0@meta.data$Status = factor(findclust_0@meta.data$Status, levels = c('NRM','M','D','E','NF','F'))
dat2 = findclust_0@meta.data%>%
  group_by(individual, Status, RNA_snn_res.0.1)%>%
  summarize(mean_cyto = mean(cyto),
            se_cyto =sd(cyto)/sqrt(n()))
```

```{r}
ggplot(dat2, aes(x = RNA_snn_res.0.1, y = mean_cyto, color = Status))+
  geom_boxplot()
```
I think findsubclusters is objectively better

```{r}
allcell_findclust_0 = findclust_0@meta.data%>%
  group_by(individual, Status)%>%
  summarize(total_cells = n())

subcell_findclust_0 = findclust_0@meta.data%>%
  group_by(individual, Status, RNA_snn_res.0.1)%>%
  summarize(cells = n())

prop_cells_findclust0 = subcell_findclust_0%>%
  right_join(allcell_findclust_0, by = 'individual')

prop_cells_findclust0%>%glimpse()
```
```{r}
ggplot(prop_cells_findclust0, aes(x = RNA_snn_res.0.1, y = cells/total_cells, color = Status.x))+
  geom_boxplot()
```
Yeah I mean cmon

I think part of it might be the lack of chromatin data, the wsnn is necessary

```{r}

```


