# Exploring genes regulating DHEA

Libraries

```{r}
library(Seurat)
library(tidyverse)
library(lme4)
library(car)
library(emmeans)
mean_expression_plot<- readRDS('Functions/mean_expression_plot.rds')

```

Read in obj

```{r}
obj <- readRDS("C:/Users/Gabe/Desktop/RNA Object.rds")

```

```{r}
human_to_ocellaris <- read_csv("C:/Users/Gabe/Desktop/multiome_poa/Reference/hsapiens_to_aocellaris.csv")

```

Genes to investigate

cyp11a1, cyp17a1, cyp11b, cyp11c1, cyp19a1b, hsd3b1, hsd17b1, hsd17b3, hsd11b2, SRD5a

Cyp11a1

```{r}
mean_expression_plot(obj, 'LOC111587899' # cyp11a1
                     )
```

Steadily increases; cyp11a1 catalyzes cholesterol to pregnenolone suggesting increased de novo steroid synthesis in females

cyp17a1

mean_expression_plot(obj, 'LOC111563640' \# cyp17a1

)

```{r}
mean_expression_plot(obj, 'LOC111563640' # cyp17a1
                     )

```

Cyp17a1 steadily decreases, with large variation in dominants.

Cyp17a1 catalyzes:

DHEA - Androstenedione

Pregnenolone - 17a hydroxyprenenolone

Progesterone - 17Oh progesterone

17 oh progesterone - androstenedione

The idea would be decreased androstenedione synthesis I guess?

I cant find any evidence of cyp11c or b so...

Brain aromatase

```{r}
mean_expression_plot(obj, 'LOC111577263'  # brain aromatase
                     )

```

Decreased estrogen production in dominants

hsd3b1

```{r}
mean_expression_plot(obj, 'hsd3b1'
                     )

```

Progressive increase in hsd3b1:

hsd3b1 catalyzes

17a-OH pregnenolone - 17a-oh-progesterone

Pregnenolone - Progesterone

Androstenedione - Testosterone

A bunch of other stuff but the androstenedione to testotsterone is the most important maybe?

hsd17b1

```{r}
mean_expression_plot(obj, 'hsd17b1'
                     )

```

Nonexistent... how?

hsd17b2

```{r}
#mean_expression_plot(obj, 'hsd17b2'
                    # )
```

Cant find any evidence it exists

hsd17b3

```{r}
mean_expression_plot(obj, 'hsd17b3'
                     )

```

Slight decrease in females

hsd17b3 catalyzes:

Androstenedione - Testosterone

11ketoandrosterone - 11KT

srd5a1

```{r}
mean_expression_plot(obj, 'srd5a1'
                     )

```

Honestly same across sexes

srd5a2

```{r}
mean_expression_plot(obj, 'LOC111564580'
                     )

```

Ok fascinating; srd5a2 catalyzes Testosterone to DHT... why is it higher in females?

The techincal name for this gene is

"3-oxo-5-alpha-steroid 4-dehydrogenase 2 - like", but that is another name for srd5a so I am confused. BUT, it has way lower mean expression than srd5a2b so that one is probably more relevant

```{r}
mean_expression_plot(obj, 'srd5a2b'
                     )
```

But this sr5a2b increases in dominants... what's the deal here?

sts

```{r}
mean_expression_plot(obj, 'sts'
                     )
```

hsd17b14

```{r}
mean_expression_plot(obj, 'hsd17b14'
                     )
```

hsd11b2 makes 11-ketoandrosterone and 11-kt

```{r}
mean_expression_plot(obj, 'hsd11b2'
                     )

```

cyp21a2

```{r}
mean_expression_plot(obj, 'LOC111588846'
                     )
```
