Here, I am going to explore a few different trajectory analyses 

Load libraries
```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(Seurat)
library(ggplot2)
library(tidyverse)

```

Read in multiome object
```{r}
obj <- readRDS("C:/Users/Gabe/Desktop/RNA Object.rds")
```
Looks good to me
```{r}
Idents(obj) <-'harmony.wnn_res0.4_clusters'
DimPlot(obj, reduction = 'harmony_wnn.umap')
```
#Diffusion based Pseudotime
```{r}
#library(devtools)
#devtools::install_git('https://github.com/theislab/destiny.git', force = T)
library(destiny)
```

Convert object to single cell experiment
```{r}
library(SingleCellExperiment)
sce <- as.SingleCellExperiment(obj)
```

Run diffusion on variable genes
```{r}
var_genes = VariableFeatures(obj)
```

Explicity specify destiny package
```{r}
dm = DiffusionMap(sce, k =30, n_pcs =20, verbose =T)
```

Extract first two diffusion components
```{r}
dc <- eigenvectors(dm)[,1:2]
```

Add coordinates to Seurat obj
```{r}
obj[['diffmap']]<- CreateDimReducObject(
  embeddings = dc,
  key = 'DC_',
  assay = 'RNA'
)
```

Calculate diffusion Pseudotime - set radial glia to root
```{r}
root_cell_type <- '2'
#find radial glia indicies
root_cell_indicies <- which(obj$harmony.wnn_res0.4_clusters == root_cell_type)

#choose median radial glia as root
root_cell_dc <- dc[root_cell_indicies,]
root_cell_center <- colMeans(root_cell_dc)

#calculate distances for all cells
dists <- apply(root_cell_dc, 1, function(x) sum((x- root_cell_center)^2))

#set smallest distance as root cell
root_cell <- root_cell_indicies[which.min(dists)]

```
Calculate diffusion pseudotime - run
```{r}
dpt <- DPT(dm, tips = root_cell)
```
Add to Seurat Object
```{r}
obj$diffusion_pseudotime <- rank(dpt$dpt) #why do rank also what is rank 
obj$diffusion_pseudotime <- 100* (obj$diffusion_pseudotime-
  min(obj$diffusion_pseudotime))/
  (max(obj$diffusion_pseudotime)-
     min(obj$diffusion_pseudotime)) ##have no idea what this does, I think it normalizes?
# ok so its 100 * (score above min)/range(score) so its a percentage I think
```

Visualize
```{r}
p1 <- ggplot(data.frame(DC1 = dc[,1],
                        DC2 = dc[,2],
                        CellType = obj$harmony.wnn_res0.4_clusters))+
  geom_point(aes(x = DC1, y = DC2, color = CellType), size = 1)+
  theme_minimal()+
  guides(color = guide_legend(override.aes =list(size =3)))+ #waht does this do??
  labs(title= "Diffusion Map", x = 'DC1',y = 'Dc2')

p2 <- FeaturePlot(obj, features = 'diffusion_pseudotime',
                  reduction = 'harmony_wnn.umap',
                  label = T,
                  pt.size = 1)+
  scale_color_viridis_c()+
  labs(title = 'Diffusion Pseudotime')

p1+p2
```

Alright I want to plot by celltype
```{r}
diff_score_data <- data.frame(individual = obj$individual,
                              Status = obj$Status, 
                              cluster = obj$harmony.wnn_res0.4_clusters,
                              pseudotime = obj$diffusion_pseudotime)%>%
  group_by(individual, Status, cluster)%>%
  summarize(pseudotime = mean(pseudotime),
            se_pseudotime = sd(pseudotime)/sqrt(n()))
library(forcats)
diff_score_data$Status <- factor(diff_score_data$Status, levels = c('NRM','M','D',"E",'NF','F'))

plot <- ggplot(diff_score_data, aes(x = fct_reorder(cluster, pseudotime), y = pseudotime))+
  geom_jitter(size = 1, alpha =0.5, aes(color = Status))+
  geom_boxplot(fill = NA, outlier.shape = NA)+
  theme_minimal()+
  labs(x = 'Cluster',y = 'Mean +/- SE Diff Pseudotime')
plot


```
Not surprising that 22 is closest because ependymal cells are related to radial glia. What is surprising is that 28 is so low on this, also 19 
Very surprising to me that OPCs and oligodendrocytes are so far on this, what? 
26 and 14, 29 should not be in this, that is probably confusing it.
29 is so variable it makes me very interested to see it's cytotrace score
It's interesting that it thinks 27 is so old, that disagrees with the CytoTRACE
It looks like there is kind of a plateau in 19-0

I think next is to look at Slingshot, that may be more accurate. My prior beleifs are that OPCs and 27 should be low on pseudotime due to their stem-nature
















